---
title: "DPLYR Crash Course"
author: "Blake Augustine"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packageCheck, include=FALSE}
mypacks <- c("ggplot2","dplyr", "quantmod", "tidyverse")  # what packages are needed?
packs <- installed.packages()   # find installed package list
install.me <- mypacks[!(mypacks %in% packs[,"Package"])]  #what needs to be installed?
if (length(install.me) >= 1) install.packages(install.me, repos = "http://cran.us.r-project.org")   # install (if needed)
lapply(mypacks, library, character.only=TRUE)  # load all packages
```

**Basics of DPLYR:**
- DPLYR is used to cut down data sets to criteria you are looking for

Basic commands to use:

`%>%`
- "piping" operator, basically indicates sequential application of dplyr functions
- x %>% y %>% z means you take x, then do y, then do z all in sequence
- %>% can be used for other functions besides dplyr ones
  + you can pipe in a ggplot() line, summary(), mean(), print(), and others!

`mutate()`
- Creates new variables in an existing data set by applying a function to a column

```{r}
# Start by creating a basic data frame
object1 <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
object2 <- c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100)
dataframe <- as.data.frame(cbind(object1, object2)) # cbind means 'column bind', combines two columns or vectors into a data frame
```

```{r}
# Use mutate() to create an object for the squared values of object1
# mutate("new object name" = function of existing object)
dataframe2 <- dataframe %>% mutate(objsqr = object1^2)
dataframe2
# Plotting the new mutated variable
ggplot(dataframe2, aes(x = object1, y = object1)) + geom_line(color = "blue", size = 1) + geom_line(aes(x = object1, y = objsqr), color = "red", size = 1)
```

`select()`
- Selects specific variables in a data set, then returns a new data set with those selected ones
- Lets look at stock data to try this:

```{r}
# Start by creating a data frame for stock data
symbols <- "PYPL"
getSymbols(symbols, from = "2020-01-01", to = "2020-12-31", src = "yahoo")
PYPL <- as.data.frame(PYPL)
head(PYPL)
```

```{r}
# Now lets select some variables:
PYPL <- PYPL %>% select(PYPL.Adjusted, PYPL.Volume)
PYPL
# Now we only have the two selected variables
```

`filter()`
- Filters the observations for a specific criteria

```{r}
# Let's filter for days with high volume:
PYPL <- PYPL %>% filter(PYPL.Volume > 10000000)
PYPL
# Now we have the same columns, but only rows where the volume is greater than 10000000
```

`slice()`
- Cuts the data set
- Usually just to get a small subset of the data to display

```{r}
# Slice the first five observations
PYPL <- PYPL %>% slice(1:5)
PYPL
# Now we have the first 5 observations
```

- After all of our commands, we are left with a data set of 2 variables with the most recent 5 observations containing high volume 
- There are more you can look into but these are probably the most commonly used!






























