---
title: "PIS Day 2 - Financial Statements and IPOs"
author: "Blake Augustine"
date: "11/15/2021"
output: 
  ioslides_presentation:
    incremental: true
    widescreen: true
    keep_md: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packageCheck, include=FALSE}
# This code chunk is a useful way to install and load packages at the same time
mypacks <- c("ggplot2","dplyr", "quantmod", "tidyverse", "finreportr", "simfinR")  # Put any packages needed in this vector
packs <- installed.packages()   
install.me <- mypacks[!(mypacks %in% packs[,"Package"])]  
if (length(install.me) >= 1) install.packages(install.me, repos = "http://cran.us.r-project.org")  
lapply(mypacks, library, character.only=TRUE)  
```

## Basics of Financial Statements | **Types** {.build}
- Balance Sheet:
  + Reports the financial position (amount of assets, liabilities, and stockholders’ equity) of an accounting entity at a point in time.
  
- Statement of Cash Flow
  + reports inflows and outflows of cash during the accounting period in the categories of operating, investing, and financing.
  
- Income Statement
  + reports the revenues less the expenses during the accounting period.

- There is also a statement of stockholder's equity, but we will be focusing on these three since we can pull them easily into R

## Basics of Financial Statements | **Balance Sheet** {.build}
- The Balance Sheet is a financial snapshot at a specific point in time.

- Assets:
  + Cash, Short-Term Investments, Accounts Receivable, Notes Receivable, Inventory (to be sold), Supplies, Prepaid Expenses, Long-Term Investments,         Equipment, Buildings, Land, Intangibles

- Liabilities: 
  + Accounts Payable, Accrued Expenses, Notes Payable, Taxes Payable, Unearned Revenue, Bonds Payable

- Owner's Equity: 
  + Common Stock, Retained Earnings
  
- Assets - Liabilities = Owner's Equity
  + We want Assets > Liabilities (most of the time)

## Basics of Financial Statements | **Income Statement** {.build}
- The Income Statement is a measure of performance of the business. 

- Revenues:
  + Cash and promises received from delivery of goods and services.
  + Examples: Sales Revenue, Fee Revenue, Interest Revenue, Rent Revenue

- Expenses:
  + Resources used to earn period’s revenues.
  + Examples: Cost of Goods Sold, Wages Expense, Rent Expense, Depreciation Expense, Insurance Expense, Repair Expense, Income Tax Expense
  
- Revenues - Expenses = Net Income
  + We want Revenue > Expenses
  
## Basics of Financial Statements | **Statement of Cash Flows** {.build}
- The Statement of Cash Flows reports inflows and outflows of cash during the accounting period.

- Cash Flows from Operating Activities:
  + Cash collected from customers less cash paid for operating expenses such as cash paid to suppliers and employees.
  
- Cash Flows from Investing Activities
  + Cash flows related to acquisition or sale of the company’s plant and equipment and investments.

- Cash Flows from Financing Activities
  + Cash flows from the receipt or payment of money to investors and creditors (except suppliers)

_______________________________________________________________________________________________________________________________________________________

## Financial Statements in R | **Balance Sheet** {.build}
```{r}
available <- simfinR_get_available_companies("WAaNgczJ0jW3urxYWXAr4TyujkEFVFvk")
```

```{r}
GOOG_statements <- simfinR_get_available_statements("18", "WAaNgczJ0jW3urxYWXAr4TyujkEFVFvk", silent = TRUE)
# pl = income statement (stands for profit-loss)
# bs = balance sheet
# cf = cash flow
GOOG_statements
```

```{r}
id_companies <- 18
api_key <- "WAaNgczJ0jW3urxYWXAr4TyujkEFVFvk"

google_bs <- simfinR_get_single_fin_statement(id_sim = id_companies, type_statement = 'bs',
period_in = 'Q3', year = 2021, api_key=api_key,
cache_folder = 'simfim_cache')
google_bs <- google_bs %>% select(acc_name, acc_value) %>% as.data.frame()
google_bs
```

## Financial Statements in R | **Income Statement** {.build}
```{r}
id_companies <- 18
api_key <- "WAaNgczJ0jW3urxYWXAr4TyujkEFVFvk"

google_pl <- simfinR_get_single_fin_statement(id_sim = id_companies, type_statement = 'pl',
period_in = 'Q3', year = 2021, api_key=api_key,
cache_folder = 'simfim_cache')
google_pl <- google_pl %>% select(acc_name, acc_value)
google_pl
```

## Financial Statements in R | **Statement of Cash Flows** {.build}
```{r}
id_companies <- 18
api_key <- "WAaNgczJ0jW3urxYWXAr4TyujkEFVFvk"

google_cf <- simfinR_get_single_fin_statement(id_sim = id_companies, type_statement = 'cf',
period_in = 'Q3', year = 2021, api_key=api_key,
cache_folder = 'simfim_cache')
google_cf <- google_cf  %>% select(acc_name, acc_value)
google_cf
```

## Accounting Ratios | **The "Important" Ones!** {.build}
- Remember when comparing ratios, compare them to other firms in the same industry

- Working Capital Ratio:
  + Company's ability to pay its current liabilities with its current assets
  + Divide current assets by current liabilities.
  + Different than total assets/liabilities!
```{r}
WC_ratio <- google_bs[21,2] / google_bs[57,2] # Accesses the data set google_bs, selects row 21, column 2, then row 57, column 2
WC_ratio
```

- Quick Ratio:
  + how well current liabilities are covered by cash and by items with a ready cash value
  + (current assets - inventory) / current liabilities
```{r}
quick_ratio <- (google_bs[21,2] - google_bs[8,2]) / google_bs[57,2]
quick_ratio
```

- Earnings per Share:
  + measures net income earned on each share of a company's common stock
  + net income / shares outstanding
```{r}
EPS <- google_pl[55,2] / 722000000
EPS
```

- Price/Earnings Ratio
  + investors' assessments of those future earnings
  + Share Price / EPS
```{r}
PE_ratio <- 2969.04 / EPS
PE_ratio
```

- Debt/Equity Ratio
  + evaluate a company's financial leverage
  + <1 indicates a low risk, safe investment
  + total liabilities / total shareholders equity
```{r}
DE_ratio <- google_bs[73,2] / google_bs[84,2]
DE_ratio
```

- Return on Equity
  + return on net assets
  + Net Income / Shareholder's Equity
```{r}
(google_pl[55,2] / google_bs[84,2]) * 100
# percent %
```

_______________________________________________________________________________________________________________________________________________________

## IPOs | **Basics** {.build}

- Initial Public Offering
  + The process of offering shares of a private corporation to the public in a new stock issuance
  + Usually really good returns if bought early!
  + Financial Ratios aren't always available for them, so more robust analysis needs to be conducted

- Andrew J. Chalk and John W. Peavy, III: *Initial Public Offerings: Daily Returns, Offering Types and the Price Effect*
  + Found that IPOs generate "abnormally high returns" following issuance.
  + Returns continued a few days after issuance too
  + Returns much greater for stocks priced below $1
  
## Final Notes | {.build}
- Recently, stocks have seen higher volatility following earnings reports (both directions)
  + Could be from distrust in the market, or other reasons
- Using our $10k in reserves we could take advantage of some of these movements, but it's going to require some analysis!

## Groups! | {.build}
- Make 2 groups:
  + One for financial statements and ratio calculation
  + One for IPOs and their dates









