---
title: "PIS Day 1 - Intro to R"
author: "Blake Augustine"
date: "10/29/2021"
output: 
  ioslides_presentation:
    incremental: true
    widescreen: true
    keep_md: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packageCheck, include=FALSE}
mypacks <- c("ggplot2","dplyr", "quantmod", "forecast", "tseries", "aTSA", "fpp2", "urca", "tidyverse", "astsa", "triangle")  # what packages are needed?
packs <- installed.packages()   # find installed package list
install.me <- mypacks[!(mypacks %in% packs[,"Package"])]  #what needs to be installed?
if (length(install.me) >= 1) install.packages(install.me, repos = "http://cran.us.r-project.org")   # install (if needed)
lapply(mypacks, library, character.only=TRUE)  # load all packages
```

## Basics of R and RMarkdown
**R Chunks**
```{r}
#```{r}
#code goes here
#code will run when outputted into a pdf, Rmd, etc.
#```
```
- By adding certain parameters to the {r} area, you can show code with output, no code with output, code with no output, etc. 
    + You can also click the gear in RMarkdown to do this
- **R Markdown:** When you `knit` a .Rmd file, your output will be whatever you decide in the header.
    + This can be a pdf, .md, slideshow, html, etc.

## Objects | **Workspace** {.build}
- This is the "environment" where R objects live. 
    + See the **Environment** tab in Rstudio
- Can be data sets, objects, lists, and more.
     + We will likely focus mostly on time series data sets in this group
- Let's create a basic object:
```{r}
object <- c(1, 2, 3, "A", "B", "C", TRUE, FALSE)
# Here we have doubles, characters, and logical data types
# The character type will supersede (coerce) the other two, so our output 
# will be in the notation for character with " "
print(object)
```


## Importing stock data using quantmod | **Libraries** {.build}
- Importing libraries allows us to access more options for plots, code, data 
  analysis, etc.
```{r, eval=FALSE}
# Type this into the console (bottom left panel)
install.packages("quantmod")
# Then type this one into an R chunk at the top of your document
library(quantmod)
```
- `quantmod` package allows us to pull stock data from the internet as a 
      time series data set
- Let's try this with Paypal from 2016-2021:
    
## Pulling Paypal Data from Yahoo | **Code** {.build}
```{r, eval=FALSE}
tickers = "PYPL"
getSymbols(tickers, from = "2010-01-01", to = "XXXX-XX-XX", src = "yahoo", 
           return.class = "ts")
period = time(PYPL)
PYPL = as.data.frame(PYPL)
```
- Our object is now created


## Viewing our data set | **Two Ways** {.build}
- We can look at our new data set by clicking on it in the `environment`
- Or we can use `head()`
```{r, include=FALSE}
tickers = "PYPL"
getSymbols(tickers, from = "2010-01-01", src = "yahoo", return.class = "ts")
period = time(PYPL)
PYPL = as.data.frame(PYPL)
PYPL$Period = period
```

```{r}
head(PYPL)
tail(PYPL)
```

_______________________________________________________________________________________________________________________________________________________

## Basic Plots: Line Graph
```{r}
plot(PYPL$Period, PYPL$PYPL.Adjusted, 
     type = "l", col = "blue", 
     ylab = "Price", xlab = "Period", main = "Adjusted Prices PYPL")
legend(x = "topleft", legend = "PYPL", lty = 1, col = "blue")
```


## Basic Plots: Histogram 
```{r}
hist(PYPL$PYPL.Adjusted)
```

_______________________________________________________________________________________________________________________________________________________

## ggplot
- `ggplot2` is a library that makes plotting more in-depth
    + allows you to plot with more than 2 data points with size, color, shape, etc.
    + allows you to add regression lines, flip coordinates, and more.
```{r, echo=FALSE}
library(ggplot2)
ggplot(PYPL, aes(x = Period, y = PYPL.Adjusted)) + geom_point(color = "Blue", size=0.4, alpha = 0.75) + 
  geom_smooth(method="lm", color="Red", linetype=3) + ggtitle("Paypal Time Series")
# OLS Fit
```

_______________________________________________________________________________________________________________________________________________________

## Two Main Focuses in R
**Discounted Cash Flows (DCF)**

- Good for present calculations, or how under or overvalued a stock might be
  in the current period
  
- Some predictability, but not perfect

- We will get to this in the future

_______________________________________________________________________________________________________________________________________________________

**Time Series Forecasting**

- Good for future valuation, or where a stock price may end up in `n` periods

- Better predictability, but requires some robust understanding of the model

```{r}
PYPL_ts <- PYPL %>% summarize(PYPL.Adjusted) %>% as.ts(PYPL.Adjusted)
tail(PYPL_ts)
```

```{r}
trend.test(PYPL_ts, plot = TRUE)
# p-value < 0.05, so there is a trend to the data
```

```{r}
paypal <- window(PYPL_ts)
forecast <- holt(paypal, h=5)
forecast
# 5 period forecast, compare to tail of PYPL_ts
```

```{r}
armod <- lm(PYPL.Adjusted ~ lag(PYPL.Adjusted), data = PYPL_ts)
armod
summary(armod)
```

```{r}
#order = c(p, d, q)
ARIMA <- arima(PYPL_ts, order = c(1, 1, 1))
ARIMA
ARIMA_fit <- PYPL_ts - resid(ARIMA)
predict_ARIMA <- predict(ARIMA)
predict_ARIMA$pred[1]
predict(ARIMA,n.ahead=5)
ts.plot(PYPL_ts)
ARIMA_forecasts <- predict(ARIMA, n.ahead = 5)$pred
ARIMA_forecast_se <- predict(ARIMA, n.ahead = 5)$se

points(ARIMA_forecasts, type = "l", col = 2)

points(ARIMA_forecasts - 2*ARIMA_forecast_se, type = "l", col = 2, lty = 2)
points(ARIMA_forecasts + 2*ARIMA_forecast_se, type = "l", col = 2, lty = 2)
```

```{r}
#order = c(p, d, q)
ARIMA <- arima(PYPL_ts, order = c(1, 1, 1))
ARIMA
ARIMA_fit <- PYPL_ts - resid(ARIMA)
predict_ARIMA <- predict(ARIMA)
predict_ARIMA$pred[1]
predict(ARIMA,n.ahead=5)
ts.plot(PYPL_ts, xlim=c(1580, 1600), ylim=c(200, 280))
ARIMA_forecasts <- predict(ARIMA, n.ahead = 5)$pred
ARIMA_forecast_se <- predict(ARIMA, n.ahead = 5)$se

points(ARIMA_forecasts, type = "l", col = 2)

points(ARIMA_forecasts - 2*ARIMA_forecast_se, type = "l", col = 2, lty = 2)
points(ARIMA_forecasts + 2*ARIMA_forecast_se, type = "l", col = 2, lty = 2)
```

_______________________________________________________________________________________________________________________________________________________

**Finally, let's take a look at our portfolio!**

```{r}
portfolio <- read.csv("portfolio.csv")
print(portfolio)
```

```{r}
sectors <- portfolio %>% select(firm, shares, market_value, sector) %>% group_by(sector) %>% summarize(n = n(), prop = n/26)
sectors
# We cover 6/11 sectors defined by finviz
ggplot(sectors, aes(x = sector, y = n)) + geom_bar(stat="identity", width=1, color="white") + coord_flip()
```

_______________________________________________________________________________________________________________________________________________________

And we will leave it there for today!

"Homework": Look at portfolio and start thinking of a sector you would like to monitor each week, and keep practicing in R!

Questions?














